A hazards approach to the biometric analysis of infant mortality

### 2023-04-05 Slides/slides.Rmd
polishing

*   Bourgeois-Pichat (1951): Biometric analysis of IM (background).
*   Hazard-based alternative.
*   Theoretical considerations
*   Real-world examples.

@@

*   Two categories: 
    *   \emp{Endogenous}: *inherited*, *delivery*, etc.
    *   \emp{Exogenous}: *accidental*, *infectious diseases*, etc.
*   How to differentiate between the two categories without infornation of causes of death?

@@

1.  \emp{Endogenous} deaths only occur during the \emp{neonatal} period (0--28 days). 
2.  On a specific time scale, \emp{exogenous} infant mortality is \emp{uniformly} distributed.

@@

\tiny
\emph{Note:} $C$ is \emp{not} part of B-P's original definition.

@@

Assume 100 births, of which 90 survives infancy, ten death ages:

@@

@@

## Väterbotten 1861-1890

@@

\includegraphics[width=0.8\textwidth]{../paper2_files/figure-latex/bioume9120-1.pdf}

@@

Post-neonatal linearity is a population property, not an individual.
\end{column}

@@

*   \emp{Advantages} of the hazards method.
       *    Fits naturally into modern survival analysis, allows \emp{censored} and \emp{truncated data}, and 
       *    The \emp{exogenous} component forms a class of \emp{proportional hazards} distributions.

@@

*   \emp{Suggestions}.
    *   Work on the \emp{natural} time scale, if access to a computer.
    *   Try \emp{standard} distributions for \emp{exogenous}  mortality. 
    *   \emp{Remember} that we are dealing with \emp{competing risks}!
        *   Requires special care.

>>>>

*   Jean \emp{Bourgeois-Pichat} (1951): \emp{Biometric} analysis of infant mortality (background).
*   \emp{Hazard}-based \emp{alternative}.
*   Theoretical considerations
*   Real-world examples.

@@

*   Two categories: 
    *   \emp{Endogenous}: *inherited*, *delivery*, etc.
    *   \emp{Exogenous}: *accidental*, *infectious diseases*, etc.
*   How to \emp{differentiate} between the two categories \emp{without information} of causes of death?

@@

1.  \emp{Endogenous} deaths only occur during the \emp{neonatal} period (0--28 days). 
2.  On a specific time scale, \emp{exogenous} infant mortality is \emp{uniformly} distributed.
    *   The \emp{log-cube} time scale.

@@

$t$ is age in days.

\vspace{\baselineskip}

\tiny{
\emph{Note:} $C$ is \emp{not} part of B-P's original definition.
}

@@

Assume \emp{100 births}, of which \emp{90 survives} infancy, \emp{ten death ages}:

@@

By \emp{subtraction}: $\text{Endogenous} = \text{Total} - \text{Exogenous}$

@@

## Västerbotten 1861-1890

@@

@@

Post-neonatal \emp{linearity} is a \emp{population property}, not an individual.

@@

*   \emp{Advantages} of the hazards method.
       *    Fits naturally into modern survival analysis, allows \emp{censored} and \emp{truncated data}, and 
       *    The \emp{exogenous} component forms a class of \emp{proportional hazards} distributions.
       *    So, theoretically more appropriate.

@@

## Suggestions for future work

*   Work on the \emp{natural} time scale, if 
    *   access to a computer, and
    *   suitable software.
*   Try \emp{standard} distributions for \emp{exogenous}  mortality, eg.
    *   \emp{Weibull},
    *   \emp{Gompertz}.
*   \emp{Remember} that we are dealing with \emp{competing risks}!
    *   Requires special care.

[https://github.com/goranbrostrom/ESSHC_23](https://github.com/goranbrostrom/ESSHC_23)

**R** code for this paper and talk.
